Expression QTL hotspot dissection with mediation analysis and pleiotropy testing

### 2019-07-30 analysis/paper/hotspots-paper.Rmd
made a new Makefile-docker.makefile for use with Docker

---
title: "Expression QTL hotspot dissection with mediation analysis and pleiotropy testing"
author: "Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman"
date: "7/27/2019"
bibliography: research.bib
numberSections: true
linkReferences: true
link-citations: true
output: 
  bookdown::html_document2:
    pandoc_args: 
      - --filter
      - pandoc-crossref
---

@@

>>>>

---
title: "Expression QTL hotspot dissection with mediation analysis and pleiotropy testing"
author: "Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman"
date: "7/27/2019"
bibliography: research.bib
linkReferences: true
link-citations: true
output: bookdown::html_document2
---

@@

## Chr 5 results

### 2019-07-30 analysis/paper/methods.Rmd
made a new Makefile-docker.makefile for use with Docker

We created heatmaps to examine patterns among pleiotropy test results in each hotspot. Our analysis used both 
local and nonlocal traits.
We represented the pleiotropy test statistics as a symmetric matrix, where each row was one expression
trait and each column was an expression trait. 
Excepting the cells on the diagonal, each cell, then, contained the pleiotropy test statistic
from the analysis involving the row's expression trait and the column's expression trait.
To create a heatmap, we first performed two hierarchical clusterings of rows. 
One hierarchical clustering involved only the rows and columns containing local traits, while the other contained only 
rows and columns containing nonlocal traits.
We calculated the Euclidean distances between all row pairs (in the respective 
submatrices) before clustering. Hierarchical clustering was performed using the complete 
linkage method. Initially, each row is assigned to its own cluster. An iterative 
algorithm then combines the two most similar clusters at each step. Distances between 
clusters were computed with the Lance-Williams dissimilarity update formula for complete 
linkage. Lastly, we arranged the rows according to optimal leaf ordering [@hahsler2008getting;@dendextend]. 
We then combined the
local and nonlocal traits into a single heatmap for each hotspot. 
To achieve coloring contrasts in the heatmaps, we
set the color spectrum limits at 0 (dark blue) and 5 (yellow). Pleiotropy test statistics above 5, too, are 
represented by
yellow. We annotated rows by a collection of columns on the right-hand side of the heatmap. This binary matrix 
consists of columns that summarize mediation results. Each column represents a local trait. If a row 
corresponds to a nonlocal trait that is mediated by the local trait, then the cell is colored dark green; 
otherwise, the cell is light green. The last annotation column and the annotation row above the heatmap both 
designate local traits with dark blue and nonlocal traits with light blue.

>>>>

We created heatmaps to examine patterns among pleiotropy test results in each hotspot. Our analysis used both 
local and nonlocal traits.
We represented the pleiotropy test statistics as a symmetric matrix, where each row was one expression
trait and each column was an expression trait. 
Excepting the cells on the diagonal, each cell, then, contained the pleiotropy test statistic
from the analysis involving the row's expression trait and the column's expression trait.
To create a heatmap, we first performed two hierarchical clusterings of rows. 
One hierarchical clustering involved only the rows and columns containing local traits, while the other contained only 
rows and columns containing nonlocal traits.
We calculated the Euclidean distances between all row pairs (in the respective 
submatrices) before clustering. Hierarchical clustering was performed using the complete 
linkage method. Initially, each row is assigned to its own cluster. An iterative 
algorithm then combines the two most similar clusters at each step. Distances between 
clusters were computed with the Lance-Williams dissimilarity update formula for complete 
linkage [@lance1967general]. Lastly, we arranged the rows according to optimal leaf ordering [@hahsler2008getting;@dendextend]. 
We then combined the
local and nonlocal traits into a single heatmap for each hotspot. 
To achieve coloring contrasts in the heatmaps, we
set the color spectrum limits at 0 (dark blue) and 5 (yellow). Pleiotropy test statistics above 5, too, are 
represented by
yellow. We annotated rows by a collection of columns on the right-hand side of the heatmap. This binary matrix 
consists of columns that summarize mediation results. Each column represents a local trait. If a row 
corresponds to a nonlocal trait that is mediated by the local trait (*i.e.*, with mediation LOD difference greater than 1.5), then the cell is colored dark green; 
otherwise, the cell is light green. We didn't examine the possibility that a local trait might mediate the relationship between a QTL and another local trait; thus, all local trait rows are colored light green. The last annotation column and the annotation row above the heatmap both 
designate local traits with dark blue and nonlocal traits with light blue.
